Increasing the Upper Bound for the EvoMan Game Competition

### 2020-08-28 paper/paper.tex
Add files via upload

    \begin{abstract}
      This paper describes a comparison between algorithms for evolving agents able to play the game Evoman.
      Our team took part in the ``Evoman: Game-playing Competition for WCCI 2020'', and won second place;
      beyond finding a good agent to satisfy the requirements of the competition - which aim at a good ability to generalise -,
      we have surpassed the existing non-general, best-known upper-bound.
      We have managed to exceed this upper bound with a Proximal Policy Optimization algorithm, by discarding the competition requirements to generalise.
      We also present our other exploratory attempts: Q-learning, Genetic Algorithms, Particle Swarm Optimisation, and their PPO hybridizations. 
      Finally, we map the behaviour of our algorithm in the space of game difficulty, generating plausible extensions to the existing upper-bound.
    \end{abstract}

@@

    Our initial exploratory attempts focused on an ensemble, $2$-stage cascade method. The first stage, designed to find good candidate solutions quickly, was either a Q-Learning\cite{q_learning}, a Genetic Algorithm\cite{genetic_algorithm} or Particle Swarm Optimization\cite{pso} algorithm. The candidate solutions are weights on a fixed-structure Artificial Neural Network, which plays the Evoman game. The second stage, aimed at further refining the solutions found by the first stage, is a slower Proximal Policy Optimization\cite{ppo} algorithm.

@@

@@

    \subsection{Problem}\label{subsec:problem}
    How well an agent performs against an enemy is measured by how much damage they inflict, and how much damage to themselves they avoid - or, conversely, how much life points they have remaining, and how much life points the opponent has at the end of the game:

    \begin{gather*}
      gain = 100.01 + player\_life - enemy\_life \hspace{1cm} (1)
    \end{gather*}

@@

    \section{Approach}\label{sec:approach}
    Our initial intention was to develop a cascade ensemble method where a fast algorithm could provide a better starting point to a slower algorithm.
    We expected the classic Q-learning\cite{q_learning} algorithm to quickly obtain a decent gain. Two more exploratory algorithms we have considered are Genetic Algorithms\cite{genetic_algorithm} and Particle Swarm Optimisation\cite{pso}.
    The slower algorithm, intended to exploit the starting points provided by the first algorithm in the cascade, is a Proximal Policy Optimization\cite{ppo} algorithm.

@@

    The input and output layers are determined by the game framework. Each game frame gives us $20$ sensor values, and we also remember the past $2$ frames, alongside the movements we've made. Thus, $3 \cdot 20 + 2 = 62$ inputs.

>>>>

    \begin{abstract}
      This paper describes a comparison between algorithms for evolving agents able to play the game Evoman.
      Our team took part in the ``Evoman: Game-playing Competition for WCCI 2020'', and won second place;
      beyond finding a good agent to satisfy the requirements of the competition - which aim at a good ability to generalise -,
      we have surpassed the existing non-general, best-known upper-bound.
      We have managed to exceed this upper bound with a Proximal Policy Optimization(PPO) algorithm, by discarding the competition requirements to generalise.
      We also present our other exploratory attempts: Q-learning, Genetic Algorithms, Particle Swarm Optimisation, and their PPO hybridizations. 
      Finally, we map the behaviour of our algorithm in the space of game difficulty, generating plausible extensions to the existing upper-bound.
    \end{abstract}

@@

    Our initial exploratory attempts focused on an ensemble, $2$-stage cascade method. The first stage, designed to find good candidate solutions quickly, was either a Q-Learning\cite{q_learning}, a Genetic Algorithm\cite{genetic_algorithm} or Particle Swarm Optimization\cite{pso} algorithm. The candidate solutions are weights on a fixed-structure Artificial Neural Network, which plays the Evoman game. The second stage, aimed at further refining the solutions found by the first stage, is a slower Proximal Policy Optimization (PPO)\cite{ppo} algorithm.

@@

    All datapoints from our research are arithmetic averages of $30$ repeats.

@@

    \subsection{Problem}\label{subsec:problem}
    How well an agent performs against an enemy is measured by how much damage they inflict, and how much damage to themselves they avoid - or, conversely, how much life points they have remaining, and how much life points the opponent has at the end of the game:
    \vspace{-0.1cm}
    \begin{gather*}
      gain = 100.01 + player\_life - enemy\_life \hspace{1cm} (1)
    \end{gather*}

@@

    \section{Approach}\label{sec:approach}
    Our initial intention was to develop a cascade ensemble method where a fast algorithm could provide a better starting point to a slower algorithm.
    We expected the classic Q-learning\cite{q_learning} algorithm to quickly obtain a decent gain. Two more exploratory algorithms we have considered are Genetic Algorithms\cite{genetic_algorithm} and Particle Swarm Optimisation\cite{pso}. For our algorithmic parameters, we chose widely-used starting values, performing small-scale experiments to fine-tune them.

    The slower algorithm, intended to exploit the starting points provided by the first algorithm in the cascade, is a Proximal Policy Optimization\cite{ppo} algorithm.

@@

    The input and output layers are determined by the game framework. Each game frame gives us $20$ sensor values, and we also remember the past $2$ frames, alongside the movements we've made. Thus, $3 \cdot 20 + 2 = 62$ inputs. The ANN has $5$ outputs, corresponding to the $5$ actions accessible to the agent.
